Income and Food Price Transmission\footnote{Revised version of MS 2015310 ``Income and Price Transmission''. Submitted to \emph{Agricultural Economics} for 2. round review. I thank two anonymous reviewers for their valuable comments.}

### 2016-03-02 reply_reviewer_1.Rnw

\documentclass[12pt,english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{babel}
\usepackage{setspace}
\doublespacing
\usepackage{geometry}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=3cm,rmargin=3cm}
\usepackage{color,soul}
\usepackage[authoryear]{natbib}

\begin{document}
\SweaveOpts{concordance=TRUE}

\title{Reply to reviewer R1}
\date{\vspace{-5ex}}
\maketitle

\subsection*{Reviewer R1}
\subsubsection*{Some Formal Aspects:}
\begin{itemize}
\item \emph{Please include Abstract and Keywords on the first page.}

Done.

\item \emph{In the beginning of the paper, it would also be good to mention why this study is important. For example, why is it important to look at how price on the agricultural products transmitted differently to the different countries?}

The introduction now has more emphasis on the paper's contribution to the price transmission literature which mostly consists of single country, single commodity case studies. I also stress the policy implications of the main findings.

\item \emph{A clarification in the introduction on how authors chose the LIC, MIC or HIC groups and what database was used for such segregation (I assume World Bank database?) and actual country names by group (Appendix?) if other than standard WB methodology would be helpful.}

I have included a footnote in the analysis pointing out that I use the World Bank income classification scheme for the country groupings.

\item \emph{Literature Review section should represent literature related to the problem and suggesting why this question is relevant and what effect it has. Provide more theoretical
background as well to set the discussion in appropriate context.}

In light of this criticism I have thoroughly revised the literature review section to make it clear that its purpose is to identify factors which the literature considers important determinants of price transmission. Regarding the theoretical background, I now discuss in greater detail than before how various country characteristics are believed to affect price transmission.

\item \emph{Not clear how table 6 is constructed. Please spend some time describing the table 6 in more details.}

I have changed the table notes so that it should be clear what the numbers represent. I have also revised the main text.

\end{itemize}

\subsubsection*{Minor Comments:}
\begin{itemize}

\item \emph{Instead of rewriting equation (11), author can make just a reference to equation (1) and describe the adjustments that were made.}

I have skipped the old equation (1) instead.

\item \emph{Table 5 provides results for the panel regression. What period is this for?}

The panel regressions use monthly price data from the entire 2005-15 period. I have corrected some mistakes in the old table and revised the table notes.

\item \emph{Typos: p.2, two lines above last paragraph, double on;}

Corrected. I have rewritten the introduction to accommodate comments from both reviewers.

\item \emph{P.7, 5th line from the bottom, double domestic}

Corrected.

\item \emph{P.12, 2nd line from top, double their}

Corrected.

\item \emph{P. 16, the IMF}

Corrected. I have rewritten section 2 and 3 to accommodate comments from both reviewers.

\item \emph{P. 17 should be tables 4 and 6.}

Corrected. Note that there are only 4 tables in the revised main text.

\end{itemize}

\end{document}

### 2016-06-21 Master.tex
Initial changes in connection with AE rejection

@@

@@

<<echo=FALSE>>=
load("infl_0708_1011.Rda")
load("ts_dat.Rda")
n.full = nrow(table(ts.dat$id))
n.sub = nrow(infl)
infl = infl[infl$iso3c != "BLR", ]
@

\subsubsection{Domestic and international food prices}
Data on domestic prices are sourced from the International Labor Organization (ILO).\footnote{See \citet{ilo16}} ILO maintains a comprehensive database of domestic food price indexes (FPIs) and consumer price indexes (CPIs) that is used to monitor the cost of living and real wages. There are not many studies which draw on this database, however, despite a broad country coverage and long monthly price series. A major issue with this data is that historical series are not recalculated whenever a new base year is introduced. The same applies when other changes are made such as the inclusion or exclusion of certain items such as beverages or tobacco from the index. This means that many of the series have discrete level shifts associated with these changes. In some cases one series replace another after a period of overlap. Finally, many of the series contain missing values. To ``tidy up'' the raw data I carried out the following steps First, I restricted the sample to the period 2005:1-2014:12. Secondly, in cases where multiple complete series were available from the same country, a single one was selected. Thirdly, series with missing values were discarded. Fourthly, and most critical for the analysis below, level shifts associated with base year changes etc. were identified and removed.\footnote{Concretely I added the change in the break point from the previous value to the break point and all points succeeding it. The value of the adjusted series at the break point is thus the same as the one preceding it.} Finally, all the series were normalized with respect to a common base period (Jan 2007). In total \Sexpr{n.full} countries have complete price series for the entire ten year period considered whereas, for example, there are \Sexpr{n.sub} countries with complete price series covering the 2005-11 subperiod. There is therefore a trade-off between the number of complete price series and the length of the series.

@@

The median country experienced an annualized rate of food and headline inflation of \Sexpr{round(median(infl$fpigr.0708),1)} and \Sexpr{round(median(infl$cpigr.0708),1)} percent, respectively, in 2007-08. Nominal food prices are not only driven by exogenous world market shocks; they follow a stochastic trend together with the non-food prices, representing underlying general (core) inflation. Dividing with the CPI is a simple way of purging the FPI of underlying domestic inflation. RFPI growth, which is approximately the difference between the FPI and CPI inflation rate, was \Sexpr{round(median(infl$rfpigr.0708),1)} percent for the median country in 2007-08.\footnote{A problem with the real FPI is that domestic non-food prices are also affected by international commodity shocks. Low or even negative growth in real domestic food prices in these two periods could therefore reflect a response to the concurrent surge in non-food commodities, oil in particular, rather than a puzzling negative response to an international food price shock.} Domestic price increases were generally lower in 2010-11. The median country experienced food and headline inflation rates \Sexpr{round(median(infl$fpigr.1011),1)} and \Sexpr{round(median(infl$cpigr.1011),1)} percent and a modest RFPI growth of \Sexpr{round(median(infl$rfpigr.1011),1)} percent. There is, however, a substantial variation in the inflation rates in both periods.

@@

<<echo=FALSE>>=
library(dplyr)
load("lrm_fpi_l1_12.Rda")
fpi = select(lrm.fpi.l.1_12, -iso3c)
load("lrm_rfpi_l1_12.Rda")
rfpi = select(lrm.rfpi.l.1_12, -iso3c)
fivenum.fpi = data.frame(sapply(fpi, summary))
fivenum.rfpi = data.frame(sapply(rfpi, summary))
n = nrow(fpi)
@

The LRM is a common measure of price transmission. \cite{IMF2011}, for example, contains an analysis of food price price transmission based on a dataset of monthly FPI inflation from 31 advanced economies and 47 emerging and developing economies over the period 2000 - 2011. Their median estimated LRM pass-through of a 1 percent increase in international food inflation is 0.18 percent in developed countries and 0.34 percent in emerging and developing countries. Similarly, \citet{Ianchovichina2014} report an average estimated LRM for the eighteen Middle East and North Africa (MENA) countries of 0.25 and the majority of the countries are reported to have  food price pass-through LRMs in the order of 0.2-0.4. Both studies use 12 lags in their country specific ADL models, i.e. they assume $p=q=12$.\footnote{\cite{IMF2011} do not include exchange rates in their regressions.} My own estimates based on a sample of \Sexpr{n} countries covering the period 2005:1 to 2014:12 are very similar. The average and median FPI LRM from a model with 12 lags of all variables are \Sexpr{round(fivenum.fpi["Mean", "V12"], 2)} and \Sexpr{round(fivenum.fpi["Median", "V12"], 2)}, respectively. That being said, it is not obvious that so many lags should be included in the models. When lag length selection is based on the Akaike Information Criterion (AIC), for example, the optimal number lags in most of the models turn out to be $p=q=1$.\footnote{For each country, I calculated the AIC for each of the $12^2$ combinations of $(p,q)\, \in (1,..,12)$. Then I selected the $(p,q)$ combination with the lowest AIC.} But when a common lag length of 1 is used for all countries, average and median FPI transmission LRMs are much lower, namely \Sexpr{round(fivenum.fpi["Mean", "V1"], 2)} and \Sexpr{round(fivenum.fpi["Median", "V1"], 2)}, respectively, whereas the average of the standard errors of the LRM estimates very similar.\footnote{I also tested for residual autocorrelation with the Durbin-Watson test. There was no evidence of residual autocorrelation in either case. Standard errors of the LRMs are based on the so called Bewley transformation of the ADL model instrumented with lags of the dependent variable.} As a compromise between the long lag lengths used in the literature which might cause problems with overfitting and the short optimal lag lengths implied by the AIC criterion which leads to LRM estimates that are lower than commonly found in the literature, the country specific models are based on common lag lengths of $p=q=6$. This choice leads to average long run FPI and RFPI transmission rates of \Sexpr{round(fivenum.fpi["Mean", "V6"], 2)} and \Sexpr{round(fivenum.rfpi["Mean", "V6"], 2)}, respectively. One interpretation of the very low RFPI LRM average is that shocks to a country's FPI is passed on almost fully to its non-FPI in the long run, a phenomenon known as propagation in the literature \citep[e.g.][]{pedersen10}.

@@

<<echo=false>>=
load("tp_ctr_mean.Rda")
load("tp_ctr_mean1.Rda")
@

\subsection{Estimated income-price transmission turning points}
Table \ref{tab:tp} summarizes the estimated income turning points derived from the significant coefficients in table \ref{tab:07-08-fc-regressions}-\ref{tab:panel-regressions}. The ratios in the last column refer to the number of countries with a per capita GDP less than the estimated turning point shown in the second-last column, relative to the total number of countries in the sample with data on income. The average share of countries that are to the left of the turning point, across all specifications, is \Sexpr{round(nc.pc.mean[1],0)} percent. There are, however, important differences between the different estimators. Panel estimation, in particular, leads to a significantly higher proportion of countries to the left of the turning point than cross section estimation (\Sexpr{round(nc.pc.mean[2],0)} vs. \Sexpr{round(nc.pc.mean[3],0)} percent for the non-panel estimators) on average. In constant 2011 international dollars, the average income level in 2014 among the group of low-, lower middle- and upper middle income countries was \$1,517, \$5,752 and \$13,719, respectively \citep{worldbank16b}.\footnote{According to the World Bank classification scheme. MICs are defined as having a 2011 per capita gross national income of US\$1,026 to \$12,475, where the conversion from local currency to US dollars is based on the Atlas method.} In comparison, the income level leading to the highest predicted price transmission, averaged across all specifications, is \$\Sexpr{format(round(tp.mean[1],0), big.mark= ",", scientific= FALSE)}. Countries from the lower end of the high-income spectrum experience the highest short run food price transmission whereas lower middle income countries experience the highest long run food price transmission according to these estimates.

\begin{center}
[Table \ref{tab:tp} around here]
\par\end{center}

>>>>

\documentclass[12pt,english]{article}
\usepackage{color,soul}
\usepackage[T1]{fontenc}
\usepackage{inputenc}
\usepackage{babel}
\usepackage{url}
\usepackage{graphicx}
\usepackage[authoryear]{natbib}
\usepackage[unicode=true]{hyperref}
\usepackage[format=hang, justification=raggedright]{subfig}
\usepackage{array}
\usepackage{multirow}
\usepackage{lscape}
\usepackage{float}
\usepackage{longtable}
\usepackage{array}
\usepackage[noae]{Sweave}
\usepackage{setspace}
\doublespacing
\usepackage{geometry}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=3cm,rmargin=3cm}
\usepackage{amsmath}
\usepackage[flushleft]{threeparttable}

\begin{document}
\input{Master-concordance}

\title{Income and Food Price Transmission\footnote{Revised version of MS 2015310 ``Income and Price Transmission''. Submitted to \emph{Agricultural Economics} for 2. round review. I thank two anonymous reviewers for their valuable comments.}}

\date{\vspace{-5ex}}

\maketitle

\begin{abstract}
This paper examines determinants of international-to-domestic food price transmission using aggregate monthly price data for more than 100 countries covering the ten year period 2005-15. Findings indicate that middle-income countries experience the highest rates of food price transmission in the long run. Other determinants of food price transmission include a country's cereal import, export and consumption shares, its trade openness, the quality of its infrastructure and government involvement in grain trade. A novel model is presented which is able to explain the economic logic behind the observed inverse-U shaped relationship between income and price transmission.

\end{abstract}

\textbf{Keywords:} price transmission, food crisis, food inflation, determinants, regression analysis, income\newpage

@@

\section{Introduction}
In response to the substantial increase in international food prices during the first decade of the new millennium, a large number of recent studies have quantified global and local poverty impacts of international price shocks.\footnote{E.g. \cite{Attanasio2013,Balagtas2014,Coxhead2012,Ferreira2013,Hoyos2011,Ivanic2012,Janvry2010,Mghenyi2011,Vu2011,Wodon2010}. In addition, a supplement to volume 39 of \emph{Agricultural Economics} is entirely devoted to this topic \citep[see][]{masters08}.} A major determinant of a country's vulnerability to such shocks is the responsiveness of its domestic food prices \citep[e.g.][]{Hoyos2011}. But what are the main determinants of the domestic price responsiveness to international shocks? Are they related to trade, consumption, geography or something else? As the title of the paper suggests, a country's income level seems to be an important determinant but other factors also play a role. In particular, countries which control their international trade through an agricultural state trading enterprise are found to experience lower food price transmission on average in the short run as well as in the long run.

The literature on domestic price impacts of international commodity price shocks is vast and the topic has experienced a revival in the wake of the 2007-08 food crisis \citep[see][]{Fackler2001, kouyate16}. However, since the majority of price transmission studies focus on a single country we know little about the determinants, other than distance and borders, of price transmission differences across countries. Moreover, most price transmission studies focus on individual commodity prices rather than food prices in general. This paper therefore contributes with a cross country analysis of price transmission determinants that is based on aggregate food price data, i.e. food price indexes, rather than prices of individual agricultural commodities.

Regarding the choice between individual or aggregate prices, each approach has its strengths and weaknesses \citep{headey10}. Commodity specific analyses have a stronger microeconomic foundation in the law of one price but results do not generalize to other food goods or even to the same food good traded in other domestic markets. Transmission estimates based on aggregate price data, on the other hand, give a more comprehensive picture and therefore better reflect domestic welfare costs of international food price fluctuations. A potential weakness though of the latter approach, is that a country's food price index might not reflect the composition of poor people's food baskets which is often based mainly on grain. Another weakness is that the composition of a country's food price index differs substantially from that of the international food price indexes and as well as from other countries' price indexes, which makes it more difficult to compare transmission rates across countries. These idiosyncrasies do not, however, preclude us from studying broad patterns in price transmission rates.

The empirical analysis is based on insights derived from a simple economic model that is more suitable for aggregate price data than the law of one price workhorse model. A key prediction is that a country's income level should have an independent effect on price transmission, in addition to factors related to trade openness, geography and so on. The model predicts a novel inverse-U shaped relationship between a country's income level and its price transmission rate. This income-price transmission relationship turns out to be a robust feature of the data.

There are also other significant predictors of the domestic responsiveness to international food price shocks. Countries which import a large share of their grains and countries whose diets are mainly based on cereals are, surprisingly, found to have lower price transmission on average whereas the opposite holds for countries which export most of their grain. Countries that are more open in general in terms of their trade to GDP ratio also experience more price transmission whereas countries with an agricultural state trading enterprise experience lower price transmission. These country characteristics also significantly explain differences in domestic food inflation during the 2007-08 food crisis.

What are the policy implications of these findings? One obvious implication is that we may have to revise somewhat our perception of vulnerability and country focus in connection with international food crises. Donors and international aid organizations tend to a focus mostly on low-income food deficit countries but, according to the estimates in this paper, it seems that poor households living in middle-income countries are actually the most vulnerable to an international food price spike. Moreover, it does not seem to be the case that international price changes in general are passed on at a higher rate to countries that import most of their grains and this was not the case either during the 2007-08 food crisis.

The paper now proceeds as follows. Section \ref{sec:model} discusses the determinants of price transmission and lays out a model of a country's food price index. Section \ref{sec:analysis} presents the empirical analysis and, finally, section \ref{sec:conclusion} offers some concluding remarks.
% !Rnw root = master.Rnw

\section{Determinants of food price transmission\label{sec:model}}
What are the main determinants of a country's responsiveness to an international food price shock? I.e., which factors would we expect to influence the relationship between international and domestic food prices? To answer this question this section first reviews literature. Next, a simple model is presented of a country's food price level and its responsiveness to international shocks based on insights from the literature review.

\subsection{Literature review}
There are only a few studies carrying out actual cross section analyses of price transmission determinants. There are, however, several studies discussing this informally, especially in relation to the 2007-08 food crisis. Arguments and main findings from a selection of these studies are summarized below.

\citet{Hoyos2011} point out that domestic food inflation during the food crisis varied a lot across countries.\footnote{Domestic food inflation is defined as the percentage change in a country's Food Price Index (FPI) which is a component of the Consumer Price Index (CPI).} Furthermore, domestic food inflation was lower on average than international food inflation. \citeauthor{Hoyos2011} offer 4 reasons why was is the case: (i) weak price transmission perhaps due to poor infrastructure, (ii) differences in the composition of food baskets and the fact that domestic food baskets contain nontradables, (iii) lack of competition in the domestic food markets and (iv) government interventions for example in the form of price controls.

\citet{Abbott2009} discusses the causes and consequences of the food crisis. Regarding the cross country differences in food inflation rates, he makes three observations: i) developing countries have more basic diets which involve a greater share of staples; ii) processing and distribution margins are typically lower than in developed countries, which means that they can shrink less as costs increase. Both factors, he argues, caused a stronger domestic food inflation response among developing countries than in developed countries to the high world prices in 2007-08. And, iii) a country's exchange rate regime has a large impact on the domestic price impact of an international shock; countries which peg their currency to the dollar are more exposed to international price fluctuations than countries whose currencies follow the euro.

\citet{Lee2013} quantify the importance of various ``external'' and ``internal'' variables on domestic food inflation based on data from 72 countries over the period 2000 - 2011. The set of external variables consists of current and lagged international food inflation, as well as a set of intra- and extra-regional food inflation averages. The internal variables represent domestic supply and demand factors as well as various other country characteristics. What they find is that  countries with large food import shares have lower food inflation whereas countries experiencing large increases in their food import shares have higher food inflation on average. The latter result could, however, reflect that food imports increase when domestic domestic food prices are on the rise, as a response to food inflation rather than a cause of it. Surprisingly, the authors also find a negative relationship between a country's per capita GDP growth rate (a demand variable) and its food inflation rate. The relationship between domestic food inflation and the per capita GDP level is found to be positive albeit only marginally significant in some specifications. A depreciation of the domestic currency/US\$ exchange rate increases domestic food inflation significantly across all specifications, as expected, whereas an increase in the money supply has no effect on food inflation. More politically stable countries tend to have lower food inflation whereas countries with a higher free trade index score are found to have higher domestic food inflation.

Two studies quantify the determinants of price transmission and market integration based on individual agricultural commodity prices rather than food price indexes. \citet{kouyate16} carry out a meta-regression analysis of the effect of distance on agricultural commodity price transmission for a large number of market pairs. Their results show that geographical distance and separation by borders reduces the likelihood of price cointegration and weakens the speed of speed of price transmission significantly. These findings are consistent with the well established stylized facts from the trade literature, that borders and distance reduces trade.

Finally, \citet{greb12} consider 8 variables covering geographical, infrastructural, institutional and commodity specific factors which are believed to influence agricultural price transmission. The authors regress their own price transmission estimates, as well as estimates obtained from the literature, on these country characteristics. What they find is that the covariates explain more of the variation in the estimated speed of adjustment parameters (the $\alpha$'s) than in the cointegration parameters (the $\beta$'s). Another major finding is that more of these covariates have significant and plausible partial effects on the $\alpha$'s than on the $\beta$'s. However, none of the determinants have a robust robust impact on price transmission.

\subsection{A model of a country's food price level}
This section presents a simple model of a country's food price level that is inspired by the arguments and findings discussed above, related to determinants of domestic food inflation and price transmission. The benefit of a working with a formal economic model is that it imposes some structure on the empirical analysis. In fact, despite its simplicity, it is possible to derive several predictions from the model regarding the determinants of price transmission.

A country's consumer price index (CPI) can be expressed as a weighted average
\begin{equation}
P =a P_{f}+(1-a)P_{n}\label{eq:cpi},
\end{equation}
where $P_{f}$ and $P_{n}$ denote the food and non-food price indexes, respectively, and $a$ is the food expenditure share. Time indexes are omitted in order to avoid notational clutter. The food price index (FPI), which is of primary interest, can be written
\begin{equation}
P_{f} =b P_{c}+(1-b)P_{m}\label{eq:fpi},
\end{equation}
where $P_{c}$ and $P_{m}$ denote the farm cost and marketing components of the FPI, respectively, and $b$ is the farm cost or commodity share.\footnote{By marketing costs I refer to all other costs than those of the raw agricultural commodities included in the final consumer food price such as those associated with distribution, storage and promotion.}  Finally, the commodity component of the domestic FPI can be decomposed as 
\begin{equation}
P_{c} =c P_{w}+(1-c)P_{d}\label{eq:compi},
\end{equation}
where $P_{w}$ and $P_{d}$ denote the tradable and purely domestic subcomponents of the commodity price index, respectively, and $c$ is the share of tradable food commodities in $P_{c}$. The defining difference between tradable and domestic food commodities is that the former have prices which are determined on the world market whereas the latter prices are determined by domestic supply and demand.

The three share parameters, $a$, $b$ and $c$ represent weights on price indexes that are progressively embedded into each other. By substituting (\ref{eq:compi}) into (\ref{eq:fpi}) we get
\begin{equation}
P_{f} = b c P_{w}+R\label{eq:fpi1},
\end{equation}
where $R=b(1-c)P_{d}+(1-b)P_{m}$ is a term which does not depend on world market prices directly. With these definitions a first-order approximation to transmission of international agricultural commodity price changes to domestic consumer food prices can be written
\begin{equation}
\theta\equiv\frac{\partial P_{f}}{\partial P_{w}}=bc,\label{eq:pt}
\end{equation}
i.e., price transmission depends multiplicatively on the share of tradable commodities in the commodity price index, $c$, and the commodity cost share in the FPI, $b$. Expression (\ref{eq:pt}) is based on the assumption that there is no knock-on effect on domestic domestic prices when world market prices change, as would be appropriate in the short term. If we relax this assumption and allow domestic prices to respond to international price changes through substitution effects we would have to add the term $b(1-c)\frac{\mathrm{d}P_{d}}{\mathrm{d}P_{w}}$. For simplicity, I will ignore this second order effect in the analysis below.

By imposing some structure on the two cost shares, $b$ and $c$ we will be able to analyze determinants of a country's price transmission. In particular, we can write these shares as functions of certain country characteristics:
\begin{equation}
b=b(x_{b}), \quad c=c(x_{c}).
\end{equation}
But what are these characteristics? As I will argue below, a country's income level should feature prominently among them.

We can reasonably assume that the wealthier the country, the lower the commodity cost share in the FPI, c.f. \citet{Abbott2009}. Formally, if $y$ denotes income then $y\in x_{b}$ and $\frac{\partial b}{\partial y} <0$. Furthermore, if we believe that countries become more integrated with the world market as they develop, not only due to better infrastructure and more open markets, but also due to for example higher quality standards associated with the adoption of modern production and management practices, then it would be appropriate to also assume $y\in x_{c}$ and $\frac{\partial c}{\partial y} >0$.\footnote{We must also impose the following set of restrictions, $\lim_{y\rightarrow \infty} b(y)=0$, $\lim_{y\rightarrow \infty} c(y)=1$, $\lim_{y\rightarrow 0^{+}} b(y)=1$ and $\lim_{y\rightarrow 0~{+}} c(y)=0$, since otherwise these functions could not represent cost shares.} These assumptions in combination imply an inverse-U shaped relationship between price transmission, $\theta$, and income.

Income is not the only determinant of price transmission of course. A country's openness to trade, the quality of its infrastructure and the composition of its food consumption set are all factors which we would expect to influence the transmission of world market price fluctuations to domestic food markets. Economies that are more open in general, perhaps for historical reasons, should have a higher tradables share and therefore experience more price transmission. Similarly with countries that, due to climatic or geographical reasons, are forced to import most of their food supply. On the other hand, there are countries that are self-sufficient due to poor infrastructure, perhaps in combination with landlockedness, resulting in prohibitive transport costs. The tradables share in these countries should be very low and so should the price transmission rate. Consumption patterns are influenced by cultural as well as geographical factors. In particular, some countries consume more grains than others and the commodity component in the food price index should therefore be larger. Consequently, domestic aggregate food prices are more exposed international shocks in such countries. Finally, there are policies that are capable of affecting both shares. Trade policies in particular have a direct effect on the tradability of a good.

To sum up, there are reasons to expect a domestic FPI reaction to price fluctuations on the international agricultural commodity markets which depends non-monotonically on income. Specifically, the simple model sketched above predicts an inverse-U shaped relationship between income and price transmission. Factors related to trade openness, consumption, geography and the quality of infrastructure also, arguably, affect price transmission. 
% !Rnw root = master.Rnw

@@

\subsubsection{Domestic and international food prices}
Data on domestic prices are sourced from the International Labor Organization (ILO).\footnote{See \citet{ilo16}} ILO maintains a comprehensive database of domestic food price indexes (FPIs) and consumer price indexes (CPIs) that is used to monitor the cost of living and real wages. There are not many studies which draw on this database, however, despite a broad country coverage and long monthly price series. A major issue with this data is that historical series are not recalculated whenever a new base year is introduced. The same applies when other changes are made such as the inclusion or exclusion of certain items such as beverages or tobacco from the index. This means that many of the series have discrete level shifts associated with these changes. In some cases one series replace another after a period of overlap. Finally, many of the series contain missing values. To ``tidy up'' the raw data I carried out the following steps First, I restricted the sample to the period 2005:1-2014:12. Secondly, in cases where multiple complete series were available from the same country, a single one was selected. Thirdly, series with missing values were discarded. Fourthly, and most critical for the analysis below, level shifts associated with base year changes etc. were identified and removed.\footnote{Concretely I added the change in the break point from the previous value to the break point and all points succeeding it. The value of the adjusted series at the break point is thus the same as the one preceding it.} Finally, all the series were normalized with respect to a common base period (Jan 2007). In total 118 countries have complete price series for the entire ten year period considered whereas, for example, there are 145 countries with complete price series covering the 2005-11 subperiod. There is therefore a trade-off between the number of complete price series and the length of the series.

@@

The median country experienced an annualized rate of food and headline inflation of 9.7 and 7.1 percent, respectively, in 2007-08. Nominal food prices are not only driven by exogenous world market shocks; they follow a stochastic trend together with the non-food prices, representing underlying general (core) inflation. Dividing with the CPI is a simple way of purging the FPI of underlying domestic inflation. RFPI growth, which is approximately the difference between the FPI and CPI inflation rate, was 2.5 percent for the median country in 2007-08.\footnote{A problem with the real FPI is that domestic non-food prices are also affected by international commodity shocks. Low or even negative growth in real domestic food prices in these two periods could therefore reflect a response to the concurrent surge in non-food commodities, oil in particular, rather than a puzzling negative response to an international food price shock.} Domestic price increases were generally lower in 2010-11. The median country experienced food and headline inflation rates 3.9 and 3.5 percent and a modest RFPI growth of 0.8 percent. There is, however, a substantial variation in the inflation rates in both periods.

@@

The LRM is a common measure of price transmission. \cite{IMF2011}, for example, contains an analysis of food price price transmission based on a dataset of monthly FPI inflation from 31 advanced economies and 47 emerging and developing economies over the period 2000 - 2011. Their median estimated LRM pass-through of a 1 percent increase in international food inflation is 0.18 percent in developed countries and 0.34 percent in emerging and developing countries. Similarly, \citet{Ianchovichina2014} report an average estimated LRM for the eighteen Middle East and North Africa (MENA) countries of 0.25 and the majority of the countries are reported to have  food price pass-through LRMs in the order of 0.2-0.4. Both studies use 12 lags in their country specific ADL models, i.e. they assume $p=q=12$.\footnote{\cite{IMF2011} do not include exchange rates in their regressions.} My own estimates based on a sample of 118 countries covering the period 2005:1 to 2014:12 are very similar. The average and median FPI LRM from a model with 12 lags of all variables are 0.32 and 0.22, respectively. That being said, it is not obvious that so many lags should be included in the models. When lag length selection is based on the Akaike Information Criterion (AIC), for example, the optimal number lags in most of the models turn out to be $p=q=1$.\footnote{For each country, I calculated the AIC for each of the $12^2$ combinations of $(p,q)\, \in (1,..,12)$. Then I selected the $(p,q)$ combination with the lowest AIC.} But when a common lag length of 1 is used for all countries, average and median FPI transmission LRMs are much lower, namely 0.07 and 0.03, respectively, whereas the average of the standard errors of the LRM estimates very similar.\footnote{I also tested for residual autocorrelation with the Durbin-Watson test. There was no evidence of residual autocorrelation in either case. Standard errors of the LRMs are based on the so called Bewley transformation of the ADL model instrumented with lags of the dependent variable.} As a compromise between the long lag lengths used in the literature which might cause problems with overfitting and the short optimal lag lengths implied by the AIC criterion which leads to LRM estimates that are lower than commonly found in the literature, the country specific models are based on common lag lengths of $p=q=6$. This choice leads to average long run FPI and RFPI transmission rates of 0.14 and 0.03, respectively. One interpretation of the very low RFPI LRM average is that shocks to a country's FPI is passed on almost fully to its non-FPI in the long run, a phenomenon known as propagation in the literature \citep[e.g.][]{pedersen10}.

@@

### 2016-06-21 analysis.Rnw
Revised analysis section for 2. round rev. of paper
Initial changes in connection with AE rejection

<<echo=FALSE>>=
load("infl_0708_1011.Rda")
load("ts_dat.Rda")
n.full = nrow(table(ts.dat$id))
n.sub = nrow(infl)
infl = infl[infl$iso3c != "BLR", ]
@

@@

<<echo=FALSE>>=
library(dplyr)
load("lrm_fpi_l1_12.Rda")
fpi = select(lrm.fpi.l.1_12, -iso3c)
load("lrm_rfpi_l1_12.Rda")
rfpi = select(lrm.rfpi.l.1_12, -iso3c)
fivenum.fpi = data.frame(sapply(fpi, summary))
fivenum.rfpi = data.frame(sapply(rfpi, summary))
n = nrow(fpi)
@

@@

>>>>

<<echo=FALSE>>=
str = "C:/Users/nqw235/Google Drive/Research/pov_price_transmission/"
load(paste0(str, "infl_0708_1011.Rda"))
load(paste0(str, "ts_dat.Rda"))
n.full = nrow(table(ts.dat$id))
n.sub = nrow(infl)
infl = infl[infl$iso3c != "BLR", ]
@

@@

<<echo=FALSE>>=
library(dplyr)
load(paste0(str, "lrm_fpi_l1_12.Rda"))
fpi = select(lrm.fpi.l.1_12, -iso3c)
load(paste0(str, "lrm_rfpi_l1_12.Rda"))
rfpi = select(lrm.rfpi.l.1_12, -iso3c)
fivenum.fpi = data.frame(sapply(fpi, summary))
fivenum.rfpi = data.frame(sapply(rfpi, summary))
n = nrow(fpi)
@

@@

\begin{center}
[Table \ref{tab:tp} around here]
\par\end{center}
